Linear analysis of the vertical shear instability in astrophysical disks

### 2015-01-27 application.tex
drafted summary

>>>>

\subsubsection{Vertically-depedent thermal relaxation} 
Note that for a vertically isothermal disk and opacity $\kappa_d(T)$, 
we can restore the $z$ dependence in $\beta$ (due to the $\rho$
dependence in  $\eta$, see Eq. \ref{eta_def}) by making the transformation
$\beta(\khat)\to\beta(\khat/\sqrt{2\pi}\hat{\rho})\equiv 
\tilde{\beta}(\khat,z)$ in  Eq. \ref{real_beta}, where
$\hat{\rho}=\rho/\rho(z=0)$.  We have have repeated the example
calculation above with $\beta\to\tilde{\beta}$. We find growth rates
were only slightly reduced at low $\khat$ and eigenfunctions are
similar.

### 2015-01-27 summary.tex
drafted summary

We applied our linear results to assess the relevance of the VSI in
protoplanetary disks, specifically the Minimum Mass Solar Nebulae (MMSN),
using disk models described in \cite{chiang10}. We find the MMSN is
unstable to the VSI at $10$s of AU, with characteristic
growth times of $10$s of orbits. However, we note the strong
temperature dependence of the physical thermal relaxation timescale
in the disk model ($\beta$ decreases as $T^{-5}$ for perturbations with small
radial scale). Thus, increasing the disk temperature can signifcantly 
shorten the typical growth times for the VSI. This is also attributed
to a larger value of $\beta_\mathrm{crit}$, since increasing $T$ also
increases the aspect-ratio $\epsilon$.

@@

\emph{Realistic thermal evolution.} We have treated the thermal
relaxation time $t_c$ as a constant for simplicity. However, a
physical expression for $t_c$ will depend on 
$z$. For example, the Newtonian cooling timescale
(\S\ref{newton_cool}) will depend on $z$ for vertically non-isothermal
disks, and through the opacity since 
$\kappa_d=\kappa_d(\rho,T)$ in general.

For vertically isothermal disks and a temperature-dependent opacity
model, as considered in \S\ref{toy_relax} and Appendix \ref{mmsn},
there is still a vertical dependence of the effective thermal relaxation time
because the radiative diffusion coefficient depends on $\rho$. We
artificially discarded this vertical dependence by 
replacing $\rho\to\Sigma/H_\mathrm{iso}$, but can restore it
by making the transformation
$\beta(\khat)\to\beta(\khat/\sqrt{2\pi}\hat{\rho})\equiv
\tilde{\beta}(\khat,z)$ in  Eq. \ref{real_beta}, where
$\hat{\rho}=\rho/\rho(z=0)$.  We have repeated the example calculation
in \S\ref{mmsn_example}  with $\beta\to\tilde{\beta}$, but found
similar results.

A proper treatment for radiative diffusion will ultimately involve
the Laplacian operator. In this case additional boundary conditions 
must be supplemented. We will consider this in a future work.

@@

>>>>

We applied our linear results to assess the applicability of the VSI in
protoplanetary disks, specifically the Minimum Mass Solar Nebulae (MMSN),
using disk models described in \cite{chiang10}. We find the MMSN is
unstable to the VSI at $10$s of AU, with characteristic
growth times of $10$s of orbits. However, we note the strong
temperature dependence of the physical thermal relaxation timescale
in the disk model ($\beta$ decreases as $T^{-5}$ for perturbations with small
radial scale). Thus, increasing the disk temperature can signifcantly 
shorten the typical growth times for the VSI. This is also attributed
to a larger value of $\beta_\mathrm{crit}$, since increasing $T$ also
increases the aspect-ratio $\epsilon$.

@@

\emph{Realistic thermal evolution.} We have treated the thermal
relaxation time $t_c$ as a constant for simplicity. However, a
physical expression for $t_c$ will depend on 
$z$. For example, the Newtonian cooling timescale
(\S\ref{newton_cool}) will depend on $z$ for vertically non-isothermal
disks, and through the opacity since 
$\kappa_d=\kappa_d(\rho,T)$ in general. 
A proper treatment for radiative diffusion, on the other hand, will
ultimately involve the Laplacian operator. In this case additional
boundary conditions must be supplemented. We will consider this in a
future work.

@@

\emph{Geometry.} Studying the VSI in the shearing box has an 
inconsistency. Namely, the equilibrium temperature in the
box only depends on height (if at all). However, it is precisely the
radial temperature (or entropy) gradient in the global disk that
gives rise to vertical shear. This is not a serious issue for
perturbations with small radial wavelengths, 
%explicitly checked by including global terms 
but will need to be resolved by adopting global disk models  
if one wishes to generalize the VSI to large-scale perturbations.
